---
title: "Gapminder | Dashboard"
author: "Mauricio Costa"
format: 
    dashboard:
        nav-buttons: 
        - icon: linkedin
          href: https://www.linkedin.com/in/mauriciocostaphd/en
        - icon: github
          href: https://github.com/mauriciompc
---

```{r}
# Load packages
pacman::p_load(tidyverse, gapminder, bslib, bsicons, shiny,
               rnaturalearth, plotly, countrycode, htmltolls,
               reactable
               )
```

```{r}
#Load the data and filter by year
gap_07 <- gapminder |> 
    filter(year == 2007)
# making some variables to use
highest_life_exp_country <- gap_07 |> 
    arrange(-lifeExp) |> 
    head(1) |> 
    pull(country)

highest_life_exp_value <- gap_07 |> 
    arrange(-lifeExp) |> 
    head(1) |> 
    pull(lifeExp) |> 
    round(1)
    
lowest_life_exp_country <- gap_07 |> 
    arrange(lifeExp) |> 
    head(1) |> 
    pull(country)

lowest_life_exp_value <- gap_07 |> 
    arrange(lifeExp) |> 
    head(1) |> 
    pull(lifeExp) |> 
    round(1)

average_life_exp_value <- gap_07 |> 
    summarise(weighted.mean = sum(lifeExp * pop)/ sum(pop)) |> 
    round(1)

```

```{r}
country_shapes <- rnaturalearth::ne_countries()
# create a worldmap to see if it's work (delete after)
#ggplot(country_shapes) + geom_sf()

# create a new variable with country code
gap_07_2 <- gap_07 |> 
    mutate(
        country_code = countrycode(country, "country.name", "iso3c")
    )

# add the values on the map
gap_map <- left_join(country_shapes, gap_07_2,
          by = c("adm0_a3" = "country_code")) |> 
    mutate(tooltip_label = paste(country,
                                 round(lifeExp, 1),
                                 sep = ": ")) |> 
    ggplot() + 
    geom_sf(aes(fill = lifeExp, text = tooltip_label)) + 
    theme_void()+ 
    theme(legend.position = "none")
# Making the map interative
gap_map_plotly <- ggplotly(gap_map, tooltip = "text")

```

```{r}
top_20_countries <- gap_07_2 |> 
    arrange(desc(lifeExp)) |> 
    head(20) |> 
    mutate(tooltip_label = paste(country,
                                 round(lifeExp, 1),
                                 sep = ": ")) |> 
    ggplot(aes(y = reorder(country, lifeExp),
               x = lifeExp,
               fill = lifeExp,
               text = tooltip_label)) + 
    geom_col() + 
    geom_text(aes(label = round(lifeExp, 1)),
              nudge_x = -10,
              color = "white") +
    labs(y = "Country",
         x = "Life Exp") + 
    theme(legend.position = "none")
top_20_countries_plotly <- ggplotly(top_20_countries, tooltip = "text")
```

# HOME

## Row 1 {height="30%"}

```{r}
value_box(
  title = "Highest Life Expectancy",
  value = highest_life_exp_value,
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg = "#518fd6"),
  h3(highest_life_exp_country)
)
```

```{r}
value_box(
  title = "Lowest Life Expectancy",
  value = lowest_life_exp_value,
  showcase = bsicons::bs_icon("thermometer"),
  theme = value_box_theme(bg = "#214773"),
  h3(lowest_life_exp_country)
)
```

```{r}
value_box(
  title = "Average Life Expectancy",
  value = average_life_exp_value,
  showcase = bsicons::bs_icon("graph-up"),
  theme = value_box_theme(bg = "#3f71ab")
)
```

## Row 2 {height="70%"}

### Column 1 {width="70%"}

```{r title = "Map of Countries by Life Expectancy"}
gap_map_plotly
```

### Column 2 {width="30%"}

```{r title = "Top 20 Countries by Life Expectancy"}
top_20_countries_plotly
```

# DOWNLOAD DATA

The data used in this dashboard is shown below and can be dowloaded as a .cvs file, or you can use direct in R `gapminder` package.

```{r}

htmltools::browsable(
    tagList(
     reactable(
      gapminder,
      elementId = "gapminder-table",
      searchable = T, # search engine in the data
      filterable = T # filter in all columns of the data
    ),
    

    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('gapminder-table')"),
  )
)


```

# ABOUT DATA

***The Gapminder Foundation*** is a non-profit venture registered in Stockholm, Sweden, that promotes sustainable global development and achievement of the United Nations Millennium Development Goals by increased use and understanding of statistics and other information about social, economic, and environmental development at local, national, and global levels.\[1\]

Gapminder was founded in 2005 by Ola Rosling, Anna Rosling Rönnlund, and Hans Rosling.\[2\] The name Gapminder was derived from the "Mind the Gap" warning messages on the London Underground.\[3\] (source: Wikipedia - https://en.wikipedia.org/wiki/Gapminder_Foundation)

A video example:

<iframe width="560" height="315" src="https://www.youtube.com/embed/QNX31t7Cees?si=3YM4SbwkZaoIXHfx" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen>

</iframe>

# ABOUT ME

Hi! My name is ***Mauricio Costa*** and I am a [Psychologist]{.underline} with a background in Psychometrics.

This is my *first* dashboard using Quarto!

I'm learning diferent ways to do some dashboards and this tool is amazing!!! In order to know me in more details, you can hit the Linkedin and Github buttons.
